---
title: "Sampling Spine Surgeons and Assigning Cases"
author: "Tyler Muffly, MD"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sampling Spine Surgeons and Assigning Cases}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tyler)
library(dplyr)
library(tidyr)
library(readxl)
library(logger)
library(stringr)
```

## Overview
This vignette demonstrates a structured workflow for managing data related to spine surgeons. The key objectives are:

* Data Loading and Cleaning: Import a dataset of spine surgeons and clean it for analysis.
* Dataset Expansion: Assign multiple insurance scenarios to each physician to simulate real-world variability.
* Case Assignment: Distribute cases among lab assistants using a priority-based splitting mechanism.
* Efficient Scaling: Highlight tools and methods to manage large-scale assignments with ease.

This vignette will guide you through a step-by-step process to achieve these goals using the `tyler` package and `tidyverse` tools.

## Step 1: Load and Prepare the Data
### Description
The dataset of spine surgeons is loaded from an Excel file. The following steps are performed:

* Filtering and Cleaning: Exclude physicians not marked for inclusion and clean up column names.
* Name Standardization: Ensure consistency by capitalizing first and last names.
* State and Time Zone Mapping: Convert state abbreviations to full names and assign appropriate time zones.
* Data Enrichment: Combine multiple fields into a single descriptive column for ease of reference.

```{r cars}
# Load the data
sample_specialists <- readxl::read_xlsx("~/Dropbox (Personal)/tyler/inst/extdata/Spine Physician List.xlsx") %>%
  dplyr::rename(
    phone_number = `Correct Number`,
    specialty_primary = taxonomies_desc,
    NPI = npi
  ) %>%
  dplyr::filter(`Include (Y/N)` == "Y") %>%
  dplyr::select(-`Include (Y/N)`, -`Reason for Exclusion`) %>%
  dplyr::mutate(
    last = stringr::str_to_title(last),
    first = stringr::str_to_title(first),
    full_name = paste("Dr.", first, last),
    addresses_state = dplyr::case_when(
      `State Code` %in% state.abb ~ state.name[match(`State Code`, state.abb)],
      TRUE ~ `State Code`
    ),
    timezone = dplyr::case_when(
      addresses_state %in% c("Maine", "New Hampshire", "Vermont", "Massachusetts", "Rhode Island",
                             "Connecticut", "New York", "New Jersey", "Delaware", "Maryland",
                             "Pennsylvania", "Virginia", "West Virginia", "North Carolina",
                             "South Carolina", "Georgia", "Florida", "Ohio", "Indiana", "Michigan") ~ "Eastern Time",
      addresses_state %in% c("Wisconsin", "Illinois", "Missouri", "North Dakota", "South Dakota",
                             "Nebraska", "Kansas", "Minnesota", "Iowa", "Kentucky", "Tennessee",
                             "Alabama", "Mississippi", "Arkansas", "Louisiana", "Texas", "Oklahoma") ~ "Central Time",
      addresses_state %in% c("Montana", "Idaho", "Wyoming", "Utah", "Colorado", "New Mexico", "Arizona") ~ "Mountain Time",
      addresses_state %in% c("California", "Nevada", "Oregon", "Washington") ~ "Pacific Time",
      addresses_state == "Hawaii" ~ "Hawaii-Aleutian Time",
      addresses_state == "Alaska" ~ "Alaska Time",
      TRUE ~ "Unknown Timezone"
    ),
    NPI = paste("NPI:", NPI),
    combined_info = paste(full_name, ",", phone_number, ",", addresses_state, ",", timezone, ",", NPI)
  ) %>%
  dplyr::relocate(combined_info, .before = everything())

glimpse(sample_specialists)
```

## Step 2: Expand Data by Insurance Scenarios
### Description
To account for variability in physician behavior under different insurance scenarios, each physician is assigned multiple insurance categories (e.g., Medicare, Medicaid, BCBS). This step is critical for simulating real-world conditions.

Key Points:
* Function Used: `phase2_sample_surgeons` from the `tyler` package.
* Input Parameters:
* `data`: The cleaned dataset of physicians.
* `insurance_types`: A vector of insurance categories.
* `seed`: Ensures reproducibility for randomized assignments.
* Output: A new dataset with multiple rows per physician, each representing a unique insurance scenario.

```{r pressure, echo=FALSE}
# Assign multiple insurance scenarios
phase2_sample_surgeons_output <- tyler::phase2_sample_surgeons(
  data = sample_specialists,
  insurance_types = c("Medicare", "Medicaid", "BCBS"),
  ensure_unique_phone_numbers = TRUE,
  output_csv_path = "data/phase_2",
  seed = 1978
) %>%
  dplyr::mutate(combined_info = paste0(combined_info, ", ", insurance))

glimpse(phase2_sample_surgeons_output)
```


## Step 3: Assign Cases to Lab Assistants
### Description
To manage workloads effectively, cases are assigned to lab assistants based on specific criteria. The process ensures a fair distribution of work and creates a systematic record of assignments.

Key Points:
* Function Used: `phase0_split_calls_to_lab_assistants_and_save_by_priority`
* Input Parameters:
* `assistant_names`: A vector of lab assistant names.
* `split_column`: The column (e.g., insurance type) used to distribute cases.
* `output_folder`: Directory where the assignments are saved.
* `seed_value`: Ensures reproducibility.
Output: Files saved for each assistant, detailing their assigned cases.

```{r}
# Define lab assistants
assistant_names <- c(
  "Amy Du", "Bret Hatzinger", "Paul Botolin", "Jasmine Hartman Budnik",
  "Miranda Manfre", "Jason Sidrak", "Daniel Stokes", "Ryan Tseng",
  "Tristan Seawalt", "Christopher Hawryluk", "Lind"
)

# Assign cases and save
tyler::phase0_split_calls_to_lab_assistants_and_save_by_priority(
  input_data_or_path = phase2_sample_surgeons_output,
  output_folder = "inst/extdata/caller_assignments",
  assistant_names = assistant_names,
  seed_value = 1978,
  complete_file_prefix = "complete_version_",
  split_file_prefix = "",
  recursive_create = TRUE,
  split_column = "insurance",
  priority_values = NULL
)
```

## Conclusion
This vignette demonstrates a streamlined workflow for managing spine surgeon data, simulating real-world scenarios, and distributing workloads efficiently. Key takeaways include:

* Data Loading and Cleaning: Standardizing and preparing datasets is critical for downstream analysis.
* Dataset Expansion: Assigning multiple insurance scenarios to physicians enables robust testing and simulations.
* Case Assignment: Distributing cases among lab assistants ensures fair and systematic workload management.

Using this workflow, researchers can effectively scale and manage large datasets, ensuring reproducibility and efficiency in their projects.
